---
title: 'Key Factors Driving Load Growth in PJM and ERCOT'
subtitle: 'https://github.com/jazpritch/KarneiKogaWargoPritchett_ENV872_EDA_FinalProject.git'
author: "Hanna Karnei, Haru Koga, Jazmine Pritchett, Jaimie Wargo" 
date: "Fall 2023"
toc: True
output: pdf_document
---

```{r setup directory and packages, include=FALSE}
library(tidyverse)
library(here)
library(dplyr)
#install.packages("readxl")
library(readxl)
library(lubridate)
library(trend)
library(stats)
library(corrplot)

getwd()
```

```{r setup theme}
# Set your ggplot theme
mytheme <- theme_gray()+
  theme(plot.title = element_text(size = 16, hjust= 0),
        axis.title = element_text(size = 13),
        legend.position = 'right')

theme_set(mytheme)
```

## Load in data

```{r ERCOT}
# Read the CSV files
x2010<- read_excel("Data/Raw/ERCOT_Source_Data/2010_ercot_hourly_load_data.xls") %>% 
  rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2011<- read_excel("Data/Raw/ERCOT_Source_Data/2011_ercot_hourly_load_data.xls") %>% 
  rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2012<- read_excel("Data/Raw/ERCOT_Source_Data/2012_ercot_hourly_load_data.xls") %>% 
  rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2013<- read_excel("Data/Raw/ERCOT_Source_Data/2013_ercot_hourly_load_data.xls") %>% 
  rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2014<- read_excel("Data/Raw/ERCOT_Source_Data/2014_ercot_hourly_load_data.xls") %>% 
  rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2015<- read_excel("Data/Raw/ERCOT_Source_Data/native_load_2015.xls") %>% 
  rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2016<- read_excel("Data/Raw/ERCOT_Source_Data/native_Load_2016.xlsx") %>% 
  rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2017<- read_excel("Data/Raw/ERCOT_Source_Data/native_Load_2017.xlsx") %>% 
  rename(Hour_Ending = "Hour Ending") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2018<- read_excel("Data/Raw/ERCOT_Source_Data/Native_Load_2018.xlsx") %>% 
  rename(Hour_Ending = "HourEnding") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2019<- read_excel("Data/Raw/ERCOT_Source_Data/Native_Load_2019.xlsx") %>% 
  rename(Hour_Ending = "HourEnding") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2020<- read_excel("Data/Raw/ERCOT_Source_Data/Native_Load_2020.xlsx") %>% 
  rename(Hour_Ending = "HourEnding") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2021<- read_excel("Data/Raw/ERCOT_Source_Data/Native_Load_2021.xlsx") %>% 
  rename(Hour_Ending = "Hour Ending") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2022<- read_excel("Data/Raw/ERCOT_Source_Data/Native_Load_2022.xlsx") %>% 
  rename(Hour_Ending = "Hour Ending") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))

# Merge the data frames based on a common column
ercot <- bind_rows(x2010, x2011, x2012, x2013, x2014, x2015, x2016, x2017, x2018, x2019, x2020, x2021, x2022)

ercot_load <- ercot %>%
  select(Hour_Ending,ERCOT)
```

```{r load data PJM, include=FALSE}
# Read the CSV files
PJM_2010<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2010.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2010 <- PJM_2010[, c(2, 7)]

PJM_2011<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2011.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2011 <- PJM_2011[, c(2, 7)]

PJM_2012<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2012.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2012 <- PJM_2012[, c(2, 7)]

PJM_2013<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2013.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2013 <- PJM_2013[, c(2, 7)]

PJM_2014<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2014.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2014 <- PJM_2014[, c(2, 7)]

PJM_2015<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2015.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2015 <- PJM_2015[, c(2, 7)]

PJM_2016<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2016.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2016 <- PJM_2016[, c(2, 7)]

PJM_2017<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2017.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2017 <- PJM_2017[, c(2, 7)]

PJM_2018<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2018.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2018 <- PJM_2018[, c(2, 7)]

PJM_2019<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2019.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2019 <- PJM_2019[, c(2, 7)]

PJM_2020<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2020.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2020 <- PJM_2020[, c(2, 7)]

PJM_2021<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2021.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2021 <- PJM_2021[, c(2, 7)]

PJM_2022<- read_csv("Data/Raw/PJM_Source_Data/PJM_MeteredHourlyLoad_2022.csv") %>% 
  rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2022 <- PJM_2022[, c(2, 7)]

#Merge the data frames based on a common column
PJM_load <- bind_rows(PJM_2010, PJM_2011, PJM_2012, PJM_2013, PJM_2014, PJM_2015, PJM_2016, PJM_2017, PJM_2018, PJM_2019, PJM_2020, PJM_2021, PJM_2022)
```

# Rationale and Research Questions

\newpage

# Dataset Information

\newpage

# Exploratory Analysis

Distribution of ERCOT hourly load data, as well as aggregated by day and month.

```{r}
# Plot hourly data 

head(ercot_load)

ggplot(ercot_load, aes(x = Hour_Ending, y = ERCOT)) +
  geom_line(color = "blue") +
  geom_smooth(method = "lm") +
  ggtitle('Hourly Electricity Load in ERCOT from 2010 to 2022') + 
  labs(x = "", y = "Load (MW)")
```

```{r}
#Aggregate data by year

ercot_load_yearly <- ercot_load %>%
  mutate(
    year = year(Hour_Ending)
  ) %>%
  group_by(year) %>%
  summarise(
    Mean_Load = mean(ERCOT, na.rm = TRUE)
  ) %>% 
  drop_na()

ercot_load_yearly
```

```{r}
#Aggregate data by month

ercot_load_monthly <- ercot_load %>%
  mutate(
    Year = year(Hour_Ending),
    Month = month(Hour_Ending)
  ) %>%
  group_by(Year, Month) %>%
  summarise(
    Mean_Load = mean(ERCOT, na.rm = TRUE)
  )

ercot_load_monthly$NewDate <- 
  as.Date(paste(ercot_load_monthly$Year, ercot_load_monthly$Month, "01", sep = "-"))

head(ercot_load_monthly)

```

```{r}

#Aggregate data by day 

ercot_load_daily<- ercot_load %>%
  mutate(
    Year = year(Hour_Ending),
    Month = month(Hour_Ending),
    Day = day (Hour_Ending)
  ) %>%
  group_by(Year, Month, Day) %>%
  summarise(
    Mean_Load = mean(ERCOT, na.rm = TRUE)
  )

ercot_load_daily$NewDate <- 
  as.Date(paste(ercot_load_daily$Year, ercot_load_daily$Month,ercot_load_daily$Day, sep = "-"))

head(ercot_load_daily)

```

```{r}
graph1 <- ggplot(ercot_load_monthly, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "blue") +
  geom_smooth(method = "lm") +
  ggtitle('Monthly Electricity Load in ERCOT from 2010 to 2022') + 
  labs(x = "", y = "Average Monhtly Load (MW)")

graph1

```

```{r}
graph2 <- ggplot(ercot_load_daily, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "blue") +
  geom_smooth(method = "lm") +
  ggtitle('Daily Electricity Load in ERCOT from 2010 to 2022') + 
  labs(x = "", y = "Average Daily Load (MW)")

graph2 
```

Distribution of PJM hourly load data, as well as aggregated by day and month.

```{r}
# Plot hourly data 

ggplot(PJM_load, aes(x = Date, y = Load_MW)) +
  geom_line(color = "black") +
  geom_smooth(method = "lm") +
  ggtitle('Hourly Electricity Load in PJM from 2010 to 2022') + 
  labs(x = "", y = "Load (MW)")
```

```{r}
#Aggregate data by month

PJM_load_monthly <- PJM_load %>%
  mutate(
    Year = year(Date),
    Month = month(Date)
  ) %>%
  group_by(Year, Month) %>%
  summarise(
    Mean_Load = mean(Load_MW, na.rm = TRUE)
  )

PJM_load_monthly$NewDate <- 
  as.Date(paste(PJM_load_monthly$Year, PJM_load_monthly$Month, "01", sep = "-"))

head(PJM_load_monthly)

#Aggregate data by day 

PJM_load_daily<- PJM_load %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Day = day(Date)
  ) %>%
  group_by(Year, Month, Day) %>%
  summarise(
    Mean_Load = mean(Load_MW, na.rm = TRUE)
  )

PJM_load_daily$NewDate <- 
  as.Date(paste(PJM_load_daily$Year, PJM_load_daily$Month,PJM_load_daily$Day, sep = "-"))

head(PJM_load_daily)

```

```{r PJM monthly graph}

PJMgraph_month<- ggplot(PJM_load_monthly, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "black") +
  geom_smooth(method = "lm") +
  ggtitle('Monthly Electricity Load in PJM from 2010 to 2022') + 
  labs(x = "", y = "Average Monhtly Load (MW)")

PJMgraph_month

```

```{r PJM daily graph}

PJMgraph_daily <- ggplot(PJM_load_daily, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "black") +
  geom_smooth(method = "lm") +
  ggtitle('Daily Electricity Load in PJM from 2010 to 2022') + 
  labs(x = "", y = "Average Daily Load (MW)")

PJMgraph_daily
```

Using a boxplot to compare electricity load distribution between ERCOT versus PJM.

```{r Boxplot}
boxplot(ercot_load, PJM_load, names = c("ERCOT", "PJM"), xlab = "Operator", ylab = "Load Distribution", main = "Electricity Load Distribution in ERCOT v. PJM 2010-2022")
```

\newpage

#Time Series Analysis The null hypothesis: Electricity load in ERCOT remained the same from 2010 to 2022. Alternative hypothesis: Electricity load in ERCOT has changed from 2010 to 2022.

```{r}
#Generate ts dataframes 

ercot_load_daily_ts <- ts(
  ercot_load_daily$Mean_Load,
  start = c(2010,1), frequency = 365
)

ercot_load_monthly_ts <- ts(
  ercot_load_daily$Mean_Load,
  start = c(2010,1),
  frequency = 12
)

```

```{r}
#11 Decompose ts dfs and plot them 

daily_decomposed <- stl(ercot_load_daily_ts, s.window="periodic")
plot(daily_decomposed)

monthly_decomposed <- stl(ercot_load_monthly_ts, s.window="periodic")
plot(monthly_decomposed)
```

```{r}
monthly_load_trend <- Kendall::SeasonalMannKendall(ercot_load_monthly_ts)
summary(monthly_load_trend)
```

The seasonal Mann-Kendall is the most appropriate test because it is designed to analyze seasonal, non-parametric data, and there is a clear seasonal trend in our data. All other tests we learned about could only be used for non-seasonal data. A positive score (361963) and tau (0.341) indicate an increasing trend in the time series data. The p-value smaller than 0.05 allows us to reject the null hypothesis. Therefore, we can conclude that electricity demand has changed over time in ERCOT.

```{r}
# Extract seasonal component from the time series 

monthly_decomposed_components <- as.data.frame(monthly_decomposed$time.series[,1:3])

monthly_decomposed_deseasoned <- 
  ercot_load_monthly_ts-monthly_decomposed_components$seasonal

# Run the Mann Kendall test on the non-seasonal data

monthly_load_trend_deseason <- Kendall::MannKendall(monthly_decomposed_deseasoned)
summary(monthly_load_trend_deseason)

```

Based on the results of the Mann Kendall test, we are able to reject the null hypothesis (p\<0.05) and identify an increasing trend in the data (tau=0.34, score=4339158). Hence, the load in ERCOT has indeed changed over time; it was not simply a seasonal change. A p-value \<0.05 in the seasonal Mann Kendall test also led us to the same conclusion.

\newpage

The null hypothesis: Electricity load in PJM remained the same from 2010 to 2022. Alternative hypothesis: Electricity load in PJM has changed from 2010 to 2022.

```{r}
#Generate ts dataframes 

PJM_load_daily_ts <- ts(
  PJM_load_daily$Mean_Load,
  start = c(2010,1), frequency = 365
)

PJM_load_monthly_ts <- ts(
  PJM_load_daily$Mean_Load,
  start = c(2010,1),
  frequency = 12
)

```

```{r}
#11 Decompose ts dfs and plot them 

daily_decomposed_PJM <- stl(PJM_load_daily_ts, s.window="periodic")
plot(daily_decomposed_PJM)

monthly_decomposed_PJM <- stl(PJM_load_monthly_ts, s.window="periodic")
plot(monthly_decomposed_PJM)
```

```{r}
monthly_load_trend_PJM <- Kendall::SeasonalMannKendall(PJM_load_monthly_ts)
summary(monthly_load_trend_PJM)
```

The seasonal Mann-Kendall is the most appropriate test because it is designed to analyze seasonal, non-parametric data, and there is a clear seasonal trend in our data. All other tests we learned about could only be used for non-seasonal data. A positive score (82902071) and tau (0.0681) indicate an increasing trend in the time series data. The p-value smaller than 0.05 allows us to reject the null hypothesis. Therefore, we can conclude that electricity demand has changed over time in PJM.

```{r}
# Extract seasonal component from the time series 

monthly_decomposed_components_PJM <- as.data.frame(monthly_decomposed_PJM$time.series[,1:3])

monthly_decomposed_deseasoned_PJM <- 
  PJM_load_monthly_ts-monthly_decomposed_components_PJM$seasonal

# Run the Mann Kendall test on the non-seasonal data

monthly_load_trend_deseason_PJM <- Kendall::MannKendall(monthly_decomposed_deseasoned_PJM)
summary(monthly_load_trend_deseason_PJM)

```

Mann Kendall test rejects the null hypothesis, which indicates that there is an increasing trend in the data (tau=0.0687, score=11896706048). This means that PJM load has changed over time from 2010 to 2022 without seasonality components. A p-value \<0.05 in the seasonal Mann Kendall test also led us to the same conclusion.

```{r}
growth_PJM <- diff(PJM_load_monthly_ts) / stats::lag(PJM_load_monthly_ts)
#growth_ERCOT <- diff(PJM_load_monthly_ts) / stats::lag(PJM_load_monthly_ts)

#plot(Date[-1], growth_PJM, type = "l", col = "blue", ylab = "Growth Rate", xlab = "Time")
#lines(Date[-1], growth_ERCOT, type = "l", col = "red")

```

# Regression Analysis

## PJM Simple Linear Regression

Simple Linear Regression

The null hypothesis: There is no change in electricity load with price in PJM from years...

Alternative hypothesis: Electricity load in PJM has changed in relation to price from years...

Real-time pricing reflects the actual market conditions, while day-ahead pricing provides insight into the expected future market conditions. Both pricing data used to understand regression analysis. Market Performance Clearing Price (\$/MWh)used reflecting the cost of the movement of load in response to regulation signals, which is related to the performance of the resource in meeting the demand.

```{r Load in price data Real-time}

# Read the CSV files
rt_2012<- read_csv("Data/Raw/PJM_Source_Data/rt_2012.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2013<- read_csv("Data/Raw/PJM_Source_Data/rt_2013.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2014<- read_csv("Data/Raw/PJM_Source_Data/rt_2014.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2015<- read_csv("Data/Raw/PJM_Source_Data/rt_2015.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2016<- read_csv("Data/Raw/PJM_Source_Data/rt_2016.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2017<- read_csv("Data/Raw/PJM_Source_Data/rt_2017.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2018<- read_csv("Data/Raw/PJM_Source_Data/rt_2018.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2019<- read_csv("Data/Raw/PJM_Source_Data/rt_2019.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2020<- read_csv("Data/Raw/PJM_Source_Data/rt_2020.csv")%>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2021<- read_csv("Data/Raw/PJM_Source_Data/rt_2021.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

rt_2022<- read_csv("Data/Raw/PJM_Source_Data/rt_2022.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = rmpcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))


#Merge the data frames based on a common column
rt_load <- bind_rows(rt_2012, rt_2013, rt_2014, rt_2015, rt_2016, rt_2017, rt_2018, rt_2019, rt_2020, rt_2021, PJM_2022)
```

```{r Load in price data Day-ahead}
# Read the CSV files
dasr_2010<- read_csv("Data/Raw/PJM_Source_Data/dasr_2010.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2011<- read_csv("Data/Raw/PJM_Source_Data/dasr_2011.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2012<- read_csv("Data/Raw/PJM_Source_Data/dasr_2012.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2013<- read_csv("Data/Raw/PJM_Source_Data/dasr_2013.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2014<- read_csv("Data/Raw/PJM_Source_Data/dasr_2014.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2015<- read_csv("Data/Raw/PJM_Source_Data/dasr_2015.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2016<- read_csv("Data/Raw/PJM_Source_Data/dasr_2016.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2017<- read_csv("Data/Raw/PJM_Source_Data/dasr_2017.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2018<- read_csv("Data/Raw/PJM_Source_Data/dasr_2018.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2019<- read_csv("Data/Raw/PJM_Source_Data/dasr_2019.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2020<- read_csv("Data/Raw/PJM_Source_Data/dasr_2020.csv")%>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2021<- read_csv("Data/Raw/PJM_Source_Data/dasr_2021.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

dasr_2022<- read_csv("Data/Raw/PJM_Source_Data/dasr_2022.csv") %>% 
  rename(Date = datetime_beginning_ept, Price = dasrmcp, Load_MW = total_pjm_rt_load_mwh) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))


#Merge the data frames based on a common column
dasr_load <- bind_rows(dasr_2010, dasr_2011, dasr_2012, dasr_2013, dasr_2014, dasr_2015, dasr_2016, dasr_2017, dasr_2018, dasr_2019, dasr_2020, dasr_2021, dasr_2022)
```

```{r PJM Load v. Price Real time}

ggplot(rt_load, aes(x= Load_MW, y= Price)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  labs(x= "Load (MW/h)", y = "Price ($)")

```

```{r PJM Load v. Price Day ahead}

ggplot(dasr_load, aes(x= Load_MW, y= Price)) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) + 
  labs(x= "Load (MW/h)", y = "Price ($)")
```

```{r}
rt.regression <- lm(data = rt_load, Price ~ Load_MW)
summary(rt.regression)
```

```{r}
cor.test(rt_load$Price, rt_load$Load_MW)
```

```{r}
dasr.regression <- lm(data = dasr_load, Price ~ Load_MW)
summary(dasr.regression)
```

```{r}
cor.test(dasr_load$Price, dasr_load$Load_MW)
```

## ERCOT Multiple Linear Regression

```{r Load in ERCOT Linear Model Data}

ercot_linear_model_data <- read.csv("Data/Raw/ERCOT_Source_Data/ERCOTLinearModelData.csv")
ercot_linear_model_data <- left_join(ercot_linear_model_data, ercot_load_yearly) %>% drop_na()

ercot_linear_model_data$percent_minority <- ercot_linear_model_data$percent_minority %>% 
  as.numeric()

ercot_linear_model_data$med_income <- as.numeric(gsub(",","",ercot_linear_model_data$med_income))

ercot_linear_model_data$population <- as.numeric(gsub(",","",ercot_linear_model_data$population))

ercot_linear_model_data_totprice <- ercot_linear_model_data %>% 
  select(-c(average_price_com, average_price_ind, average_price_res))

ercot_linear_model_data_totprice
```

```{r ERCOT corrplot}
ercot.tot.Corr <- cor(ercot_linear_model_data_totprice)
corrplot(ercot.tot.Corr, method = 'ellipse')
```

```{r}
ercot.total <- lm(data = ercot_linear_model_data_totprice, 
                       Mean_Load ~ average_price_tot + population + percent_renew +
                        med_income + percent_minority)

summary(ercot.total)
```

```{r ERCOT AIC}
ercot.AIC.total <- lm(data = ercot_linear_model_data_totprice, 
                       Mean_Load ~ average_price_tot + population + percent_renew +
                        med_income + percent_minority)

step(ercot.AIC.total)
```

| Variables                                                              | AIC Value |
|-------------------------------------------|-----------------------------|
| Mean_Load + population + percent_renew + med_income + percent_minority | 178       |
| average_price_tot + population + percent_renew + med_income            | 176.04    |
| average_price_tot + percent_renew + med_income                         | 174.26    |
| average_price_tot + percent_renew                                      | 173.58    |

```{r ERCOT Regression Summary}
ercot.totalavg.regression <- lm(data = ercot_linear_model_data_totprice, 
                       Mean_Load ~ average_price_tot + percent_renew)

summary(ercot.totalavg.regression)

```

|      Variable | Coefficients | p-value |
|--------------:|--------------|---------|
|     Intercept | 15,132.37    | 0.0224  |
| average_price | 1,976.77     | 0.0083  |
| percent_renew | 358.75       | 0.0001  |

```{r ERCOT Residuals Plots}
plot(ercot.totalavg.regression, which=1)
```

```{r}
plot(ercot.totalavg.regression, which=2)
```

\newpage

# Summary and Conclusions

\newpage

# References

\<add references here if relevant, otherwise delete this section\>
