---
title: '**Key Factors Driving Load Growth in PJM and ERCOT**'
author: "Hanna Karnei, Haru Koga, Jazmine Pritchett, Jaimie Wargo" 
output: pdf_document
toc: true
---


```{r setup directory and packages, include=FALSE}
library(tidyverse)
library(here)
library(dplyr)
#install.packages("readxl")
library(readxl)
library(lubridate)
library(trend)
library(stats)
library(corrplot)

getwd()
```

```{r setup theme, include=FALSE}
# Set your ggplot theme
mytheme <- theme_gray()+
  theme(plot.title = element_text(size = 16, hjust= 0),
        axis.title = element_text(size = 13),
        legend.position = 'right')

theme_set(mytheme)
```

# Github Repository 

https://github.com/jazpritch/KarneiKogaWargoPritchett_ENV872_EDA_FinalProject


# 1. Rationale and Research Questions 


(Include image of PJM and ERCOT map area)

Written Section (Bold questions!)

\newpage

# 2. Dataset Information and Wrangling
# 2.1 PJM Load Data
# 2.2 ERCOT Load Data 
# 2.3 Demographic Data 

\newpage

# 3. Exploratory Analysis of Data
Figure 1. Boxplot
...

# 3.1 Exploration of ERCOT
# 3.2 Exploration of PJM

# 4. Analysis 
# 4.1 Time Series Analysis 
# 4.2 Simple Linear Regression of PJM
# 4.3 Multiple Linear Regression of ERCOT
Bold each hypothesis!

# 5. Summary and Conclusions

# 6. References 

