---
title: "EDE_Time_Series_ERCOT"
output: html_document
date: "2023-11-25"
---

The null hypothesis: Electricity load in ERCOT remained the same from 2010 to 2023. 
Alternative hypothesis: Electricity load in ERCOT has changed from 2010 to 2023. 

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(dplyr)
#install.packages("readxl")
library(readxl)
library(lubridate)
library(trend)
setwd("/Users/hannakarnei/Desktop/EDA/EDE_Final_Project")
getwd()
```

```{r}
# Read the CSV files
x2010<- read_excel("ERCOT_Source_Data/2010_ercot_hourly_load_data.xls") %>% rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2011<- read_excel("ERCOT_Source_Data/2011_ercot_hourly_load_data.xls") %>% rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2012<- read_excel("ERCOT_Source_Data/2012_ercot_hourly_load_data.xls") %>% rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2013<- read_excel("ERCOT_Source_Data/2013_ercot_hourly_load_data.xls") %>% rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2014<- read_excel("ERCOT_Source_Data/2014_ercot_hourly_load_data.xls") %>% rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2015<- read_excel("ERCOT_Source_Data/native_load_2015.xls") %>% rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2016<- read_excel("ERCOT_Source_Data/native_Load_2016.xlsx") %>% rename(Hour_Ending = Hour_End) %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2017<- read_excel("ERCOT_Source_Data/native_Load_2017.xlsx") %>% rename(Hour_Ending = "Hour Ending") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2018<- read_excel("ERCOT_Source_Data/Native_Load_2018.xlsx") %>% rename(Hour_Ending = "HourEnding") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2019<- read_excel("ERCOT_Source_Data/Native_Load_2019.xlsx") %>% rename(Hour_Ending = "HourEnding") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2020<- read_excel("ERCOT_Source_Data/Native_Load_2020.xlsx") %>% rename(Hour_Ending = "HourEnding") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2021<- read_excel("ERCOT_Source_Data/Native_Load_2021.xlsx") %>% rename(Hour_Ending = "Hour Ending") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2022<- read_excel("ERCOT_Source_Data/Native_Load_2022.xlsx") %>% rename(Hour_Ending = "Hour Ending") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))
x2023<- read_excel("ERCOT_Source_Data/Native_Load_2023.xlsx") %>% rename(Hour_Ending = "Hour Ending") %>%
  mutate(Hour_Ending = as.Date(Hour_Ending, format = "%m/%d/%Y"))

# Merge the data frames based on a common column
ercot <- bind_rows(x2010, x2011, x2012, x2013, x2014, x2015, x2016, x2017, x2018, x2019, x2020, x2021, x2022, x2023)
head(ercot)

```
```{r}

# Plot hourly data 

ercot_load <- ercot %>%
  select(Hour_Ending,ERCOT)

head(ercot_load)

ggplot(ercot_load, aes(x = Hour_Ending, y = ERCOT)) +
  geom_line(color = "blue") +
  geom_smooth(method = "lm") +
  ggtitle('Hourly Electricity Load in ERCOT from 2010 to 2023') + 
  labs(x = "", y = "Load (MW)")
```

```{r}
#Aggregate data by month

ercot_load_monthly <- ercot_load %>%
  mutate(
    Year = year(Hour_Ending),
    Month = month(Hour_Ending)
  ) %>%
  group_by(Year, Month) %>%
  summarise(
    Mean_Load = mean(ERCOT, na.rm = TRUE)
  )

ercot_load_monthly$NewDate <- 
  as.Date(paste(ercot_load_monthly$Year, ercot_load_monthly$Month, "01", sep = "-"))

head(ercot_load_monthly)

```
```{r}

#Aggregate data by day 

ercot_load_daily<- ercot_load %>%
  mutate(
    Year = year(Hour_Ending),
    Month = month(Hour_Ending),
    Day = day (Hour_Ending)
  ) %>%
  group_by(Year, Month, Day) %>%
  summarise(
    Mean_Load = mean(ERCOT, na.rm = TRUE)
  )

ercot_load_daily$NewDate <- 
  as.Date(paste(ercot_load_daily$Year, ercot_load_daily$Month,ercot_load_daily$Day, sep = "-"))

head(ercot_load_daily)

```


```{r}
graph1 <- ggplot(ercot_load_monthly, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "blue") +
  geom_smooth(method = "lm") +
  ggtitle('Monthly Electricity Load in ERCOT from 2010 to 2023') + 
  labs(x = "", y = "Average Monhtly Load (MW)")

graph1

```


```{r}
graph2 <- ggplot(ercot_load_daily, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "blue") +
  geom_smooth(method = "lm") +
  ggtitle('Daily Electricity Load in ERCOT from 2010 to 2023') + 
  labs(x = "", y = "Average Daily Load (MW)")

graph2 
```

```{r}
#Generate ts dataframes 

ercot_load_daily_ts <- ts(
  ercot_load_daily$Mean_Load,
  start = c(2010,1), frequency = 365
)

ercot_load_monthly_ts <- ts(
  ercot_load_daily$Mean_Load,
  start = c(2010,1),
  frequency = 12
)

```

```{r}
#11 Decompose ts dfs and plot them 

daily_decomposed <- stl(ercot_load_daily_ts, s.window="periodic")
plot(daily_decomposed)

monthly_decomposed <- stl(ercot_load_monthly_ts, s.window="periodic")
plot(monthly_decomposed)
```
```{r}
monthly_load_trend <- Kendall::SeasonalMannKendall(ercot_load_monthly_ts)
summary(monthly_load_trend)
```

The seasonal Mann-Kendall is the most appropriate test because it is designed to analyze seasonal, non-parametric data, and there is a clear seasonal trend in our data. All other tests we learned about could only be used for non-seasonal data. A positive score (361963) and tau (0.341) indicate an increasing trend in the time series data. The p-value smaller than 0.05 allows us to reject the null hypothesis. Therefore, we can conclude that electricity demand has changed over time in ERCOT. 


```{r}
# Extract seasonal component from the time series 

monthly_decomposed_components <- as.data.frame(monthly_decomposed$time.series[,1:3])

monthly_decomposed_deseasoned <- 
  ercot_load_monthly_ts-monthly_decomposed_components$seasonal

# Run the Mann Kendall test on the non-seasonal data

monthly_load_trend_deseason <- Kendall::MannKendall(monthly_decomposed_deseasoned)
summary(monthly_load_trend_deseason)

```
Based on the results of the Mann Kendall test, we are able to reject the null hypothesis (p<0.05) and identify an increasing trend in the data (tau=0.34, score=4339158). Hence, the load in ERCOT has indeed changed over time; it was not simply a seasonal change. A p-value <0.05  in the seasonal Mann Kendall test also led us to the same conclusion. 

