---
title: "EDE_Time_Series_ERCOT"
output: html_document
date: "2023-11-25"
---

The null hypothesis: Electricity load in PJM remained the same from 2010 to 2022. 
Alternative hypothesis: Electricity load in ERCOT has changed from 2010 to 2022. 

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(dplyr)
#install.packages("readxl")
library(readxl)
library(lubridate)
library(trend)
library(stats)
getwd()
```

```{r}
# Read the CSV files
PJM_2010<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2010.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2010 <- PJM_2010[, c(2, 7)]
PJM_2011<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2011.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2011 <- PJM_2011[, c(2, 7)]
PJM_2012<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2012.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2012 <- PJM_2012[, c(2, 7)]
PJM_2013<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2013.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2013 <- PJM_2013[, c(2, 7)]
PJM_2014<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2014.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2014 <- PJM_2014[, c(2, 7)]
PJM_2015<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2015.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2015 <- PJM_2015[, c(2, 7)]
PJM_2016<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2016.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2016 <- PJM_2016[, c(2, 7)]
PJM_2017<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2017.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2017 <- PJM_2017[, c(2, 7)]
PJM_2018<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2018.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2018 <- PJM_2018[, c(2, 7)]
PJM_2019<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2019.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2019 <- PJM_2019[, c(2, 7)]
PJM_2020<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2020.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2020 <- PJM_2020[, c(2, 7)]
PJM_2021<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2021.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2021 <- PJM_2021[, c(2, 7)]
PJM_2022<- read_csv("FinalProject/PJM_MeteredHourlyLoad_2022.csv") %>% rename(Date = datetime_beginning_ept, Load_MW = mw) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))
PJM_2022 <- PJM_2022[, c(2, 7)]

# # Merge the data frames based on a common column
PJM_load <- bind_rows(PJM_2010, PJM_2011, PJM_2012, PJM_2013, PJM_2014, PJM_2015, PJM_2016, PJM_2017, PJM_2018, PJM_2019, PJM_2020, PJM_2021, PJM_2022)
head(PJM_load)

```

```{r}

# Plot hourly data 

ggplot(PJM_load, aes(x = Date, y = Load_MW)) +
  geom_line(color = "black") +
  geom_smooth(method = "lm") +
  ggtitle('Hourly Electricity Load in PJM from 2010 to 2022') + 
  labs(x = "", y = "Load (MW)")
```

```{r}
#Aggregate data by month

PJM_load_monthly <- PJM_load %>%
  mutate(
    Year = year(Date),
    Month = month(Date)
  ) %>%
  group_by(Year, Month) %>%
  summarise(
    Mean_Load = mean(Load_MW, na.rm = TRUE)
  )

PJM_load_monthly$NewDate <- 
  as.Date(paste(PJM_load_monthly$Year, PJM_load_monthly$Month, "01", sep = "-"))

head(PJM_load_monthly)

#Aggregate data by day 

PJM_load_daily<- PJM_load %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Day = day(Date)
  ) %>%
  group_by(Year, Month, Day) %>%
  summarise(
    Mean_Load = mean(Load_MW, na.rm = TRUE)
  )

PJM_load_daily$NewDate <- 
  as.Date(paste(PJM_load_daily$Year, PJM_load_daily$Month,PJM_load_daily$Day, sep = "-"))

head(PJM_load_daily)

```


```{r}
PJMgraph_month<- ggplot(PJM_load_monthly, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "black") +
  geom_smooth(method = "lm") +
  ggtitle('Monthly Electricity Load in PJM from 2010 to 2022') + 
  labs(x = "", y = "Average Monhtly Load (MW)")

graph_PJM

```


```{r}
PJMgraph_daily <- ggplot(PJM_load_daily, aes(x = NewDate, y = Mean_Load)) +
  geom_line(color = "black") +
  geom_smooth(method = "lm") +
  ggtitle('Daily Electricity Load in PJM from 2010 to 2022') + 
  labs(x = "", y = "Average Daily Load (MW)")

PJMgraph_daily
```

```{r}
#Generate ts dataframes 

PJM_load_daily_ts <- ts(
  PJM_load_daily$Mean_Load,
  start = c(2010,1), frequency = 365
)

PJM_load_monthly_ts <- ts(
  PJM_load_daily$Mean_Load,
  start = c(2010,1),
  frequency = 12
)

```

```{r}
#11 Decompose ts dfs and plot them 

daily_decomposed_PJM <- stl(PJM_load_daily_ts, s.window="periodic")
plot(daily_decomposed_PJM)

monthly_decomposed_PJM <- stl(PJM_load_monthly_ts, s.window="periodic")
plot(monthly_decomposed_PJM)
```
```{r}
monthly_load_trend_PJM <- Kendall::SeasonalMannKendall(PJM_load_monthly_ts)
summary(monthly_load_trend_PJM)
```

The seasonal Mann-Kendall is the most appropriate test because it is designed to analyze seasonal, non-parametric data, and there is a clear seasonal trend in our data. All other tests we learned about could only be used for non-seasonal data. A positive score (82902071) and tau (0.0681) indicate an increasing trend in the time series data. The p-value smaller than 0.05 allows us to reject the null hypothesis. Therefore, we can conclude that electricity demand has changed over time in PJM. 


```{r}
# Extract seasonal component from the time series 

monthly_decomposed_components_PJM <- as.data.frame(monthly_decomposed_PJM$time.series[,1:3])

monthly_decomposed_deseasoned_PJM <- 
  PJM_load_monthly_ts-monthly_decomposed_components_PJM$seasonal

# Run the Mann Kendall test on the non-seasonal data

monthly_load_trend_deseason_PJM <- Kendall::MannKendall(monthly_decomposed_deseasoned_PJM)
summary(monthly_load_trend_deseason_PJM)

```
Mann Kendall test rejects the null hypothesis, which indicates that there is an increasing trend in the data (tau=0.0687, score=11896706048). This means that PJM load has changed over time from 2010 to 2022 without seasonality components. A p-value <0.05  in the seasonal Mann Kendall test also led us to the same conclusion. 

```{r}
growth_PJM <- diff(PJM_load_monthly_ts) / stats::lag(PJM_load_monthly_ts)
#growth_ERCOT <- diff(PJM_load_monthly_ts) / stats::lag(PJM_load_monthly_ts)

#plot(Date[-1], growth_PJM, type = "l", col = "blue", ylab = "Growth Rate", xlab = "Time")
#lines(Date[-1], growth_ERCOT, type = "l", col = "red")

```

